---
title: "class 15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
metaFile <- "GSE37704_metadata.csv"
countFile <- "GSE37704_featurecounts.csv"
```


```{r}
# Import metadata and take a peak
colData = read.csv(metaFile, row.names=1)
head(colData)
```

```{r}
# Import countdata
countData = read.csv(countFile, row.names=1)
head(countData)
```

```{r}
# Note we need to remove the odd first $length col
countData <- as.matrix(countData[,-1])
head(countData)
```

```{r}
colnames(countData)
```

```{r}
rownames(colData)
```

Are all of them true?
```{r}
all (colnames(countData) == rownames(colData))
```

We want to remove genes that have 0 count values in all experiments (i.e. rows that have 0 across all columns)
```{r}
countData= countData[rowSums(countData) != 0,]
head(countData)
```

#DeSeq Analysis 
```{r}
library(DESeq2)
```

```{r}
#Set up the object with our data in the way DESeq wants it
dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
dds = DESeq(dds)
```

Get our results
```{r}
res = results(dds, contrast=c("condition", "hoxa1_kd", "control_sirna"))
```
```{r}
res
```

#Volcano Plot
```{r}
plot( res$log2FoldChange, -log(res$padj) )
```

Improve this plot by completing the below code, which adds color and axis labels
```{r}
# Make a color vector for all genes
mycols <- rep("gray", nrow(res) )

# Color red the genes with absolute fold change above 2
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"

# Color blue those with adjusted p-value less than 0.01
#  and absolute fold change more than 2
inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

plot( res$log2FoldChange, -log(res$padj), col=mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
```

```{r eval=FALSE}
BiocManager::install("AnnotationDbi")
BiocManager::install("org.Hs.eg.db")
```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)
```

```{r}
res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(countData),  #where are my IDs
                    keytype="ENSEMBL",          #what format are my IDs
                    column="SYMBOL",            #the new format I want
                    multiVals="first")
res
```

```{r}
res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(countData),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")
res$entrez
```

```{r}
library(gage)
library(gageData)
```

```{r}
data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```

The main gage() function requires a named vector of fold changes, where the names of the values are the Entrez gene IDs.
```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
# Get the results
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

lets look at the object returned from gage().
```{r}
attributes(keggres)
```

```{r}
# Look at the first few down (less) pathways
head(keggres$less)
```

Now, let's try out the pathview() function from the pathview package to make a pathway plot with our RNA-Seq expression results shown in color.
```{r}
library(pathview)
```
```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```

