---
title: "miniproject"
author: "Shilpa Nath"
date: "10/29/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Data input 

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

We have `r nrow(wisc.df)` samples in this dataset 
```{r}
nrow(wisc.df)
```

How many benign (not cancerous) and malignant (cancerous) samples do we have in the datatset?
```{r}
table(wisc.df$diagnosis)
```

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])
head(wisc.data)

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
#head(wisc.data)
```

Store the diagnosis for reference in the future as a separate vector 
```{r}
diagnosis <- wisc.df$diagnosis
```

##Questions
Q1. How many observations are in this dataset?
```{r}
nrow(wisc.df)
```

Q2. How many of the observations have a malignant diagnosis?
```{r}
table(wisc.df$diagnosis)
```

Q3. How many variables/features in the data are suffixed with _mean?
```{r}
colnames(wisc.df)
```
```{r}
grep("_mean", colnames(wisc.df), value = TRUE)
```
```{r}
length(grep("_mean", colnames(wisc.df)))
```

```{r}
round(colMeans(wisc.data), 3)
```

```{r}
round( apply(wisc.data,2,sd), 3)
```

These values look very different so I will use `scale=TRUE`
```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

```{r}
plot(wisc.pr)
```
Let's make a plot of PC1 vs. PC2
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis)
```
#Questions
Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
```{r}
x <- summary(wisc.pr)
x$importance[,1]
```

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
```{r}
which(x$importance[3,] > 0.9)[1]
```
# Create a biplot of wisc.pr
```{r}
biplot(wisc.pr)
```
Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?
A: Difficult to understand the plot above because too much data. 

# Scatter plot observations by components 1 and 2
```{r}
plot(wisc.pr$x[, c(1, 2)], col = (diagnosis), 
     xlab = "PC1", ylab = "PC2")
```

Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?
A: 

#Hierarchical Clustering 
# Scale the wisc.data data: data.scaled
```{r}
data.scaled <- scale(wisc.data)
```
# Calculate the (Euclidean) distances: data.dist
```{r}
data.dist <- dist(data.scaled)
```
# Create a hierarchical clustering model: wisc.hclust
```{r}
wisc.hclust <- hclust(data.dist, method = "complete")
```

```{r}
plot(wisc.hclust)
```



#Actual Hierarchical Clustering 
```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]), method = "complete")

plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

#Section 7 Predictions

#url <- "new_samples.csv"
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```
```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15, cex=3)
text(npc[,1], npc[,2], labels= c(1,2), col="white")
```

Q17. Which of these new patients should we prioritize for follow up based on your results?
A: 
