---
title: "class 7 r function and packages"
author: "Shilpa Nath"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R functions revisited 

Source my functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```

```{r}
rescale(c(1,10,5,NA,6))
```

write a fnx to find where the na elements in two input vectors

First make some simple input where I know the answer
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

Looked online and found the **is.na()** fnx
```{r}
is.na(x)
is.na(y)
```
The AND fnx requires two input TRUE to give a TRUE output
```{r}
is.na(x) & is.na(y)
```
Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have. This is my working snippet.
```{r}
sum(is.na(x) & is.na(y))
```

and the **which()** fnx tells me where the TRUE values are 
```{r}
which(is.na(x))
```

```{r}
which(is.na(y))
```

Now turn it into a function.
```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

Test fnx
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x,y2)
```

```{r}
x2 <- c(NA,NA)
both_na(x2,y2)
```

```{r}
x <- c(1, NA, NA)
y3 <- c( 1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```
```{r}
 both_na2 <- function(x,y) {
   if(length(x) != length(y)) {
 stop("Input x and y should be the same length")
 }
  sum(is.na(x) & is.na(y))
}
```



```{r}
x3 <- c(1,NA,NA,1,NA,NA,1)
y3 <-  c(1,NA,NA,NA,NA,NA,NA,NA)
```

#Practice
```{r}
# student 1
x <- c(100, 100, 100, 100, 100, 100, 100, 90)
# student 2
y <- c(100, NA, 90, 90, 90, 90, 97, 80)

mean(x [-which.min(x)])
```
```{r}
mean(y [-which.min(y)])
```
```{r}
which.min(y)
y[-which.min(y)]
mean(y[-which.min(y)], na.rm = TRUE)
```
We have our working code now turn it into a fnx
both_na2 <- function(x,y) {
   if(length(x) != length(y)) {
 stop("Input x and y should be the same length")
 }
  sum(is.na(x) & is.na(y))
}
```{r}
grade <- function(x,y)
  if(any(is.na(x)) ) {
    warning("Student is missing a homework")
  mean(x[-which.min(x)], na.rm=TRUE)
}
```

```{r}
grade2 <- function(x,y)
  if(any(is.na(x)) ) {
    warning("Student is missing a homework")
  sum(x[-which.min(x)], na.rm=TRUE)/length(x)-1
}
```

```{r}
url <- "https://tinyurl.com/gradeinput"
read.csv(url, row.names = 1)
hw <- read.csv(url, row.names=1)
```

```{r}
apply(hw, 1, grade)
```

```{r}
install.packages("BiocManager")
library(BiocManager)
install("rayshader")
```

```{r}

```

